Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2024-01-19 E-Hypergraphs-LICS24.tex
Figures, padding and typos

\usepackage{subcaption}

@@

\begin{figure}
\[
	\scalebox{0.65}{
	\tikzfig{categorical-semantics/egraph-strings}
	}
\]
\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
\end{figure}\label{fig:egraph-strings}

@@

\begin{remark}
    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of verticies and edges.
    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
    For example, an immediate successor of a vertex $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
\end{remark}

@@

We then quotient the set verticies and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
    \begin{align*}
        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
        & \}    
    \end{align*}
We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{i.e.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
We have 
\[
    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
\]
and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
We then define 
\[
    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}
\]
We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
Likewise, 
\[
    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
\]

@@

Recall that $<$ is essentially a transitive clousure of $<^{\mu}$ and hence we can first define 
    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
    \] and 
    \[
        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
    We will consider several cases.
    First, assume that $u$ is an element of $V_{X+Y}$ and
    \begin{enumerate}
        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
        %       The last requirement means that there necessarily exist $x = <_{c,X}^{\mu}(f_{E}(z)) = f_{E}(<_{c,Z}^{\mu}(z))$ and $y = <_{c,Y}^{\mu}(g_{E}(z)) = g_{E}(<_{c,Z}^{\mu}(z))$.
        %       We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
              The last condition means that there exists $e_y = <_{Y}^{\mu}(\iota_{V_{Y}}(g_{V}(z)))$.
              We then define 
              \[
                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(e_y)]
              \]
        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
              The one before the last condition means that there exists $e_x = <_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z)))$.
              We then define 
              \[
                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)] = [\iota_{1,E}(e_x)]
              \]
        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u)) = e'$
        \[
            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] = [e']
        \]
        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$). 
              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
              \[ 
                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[v])
              \]
    \end{enumerate}
    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
    Now, assume that $u$ is an element of $E_{X+Y}$.
    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
    Consider two cases.
    \begin{enumerate}
    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u)) = u_1$.
            Then we define
        \[
            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u))]
        \]
    \item $[u) = \varnothing$ and there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v]) = [e']$ and such that there is an undirected path from $[u]$ to $[v]$, we define 
        \[
            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
        \] 
    \end{enumerate}
    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
    % Essentially, this interleaving means that verticies inherit the relation from incident edges and vice versa.
    Clearly, all the cases above are disjoint.

@@

    Let's check that this is a well-defined function.
    Suppose $t_1, t_2$ are verticies and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
    \begin{itemize}
        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$
        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$
    \end{itemize} 
    Consider the first case. Then, by definition.
    \[
        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
    \]
    and
    \[
        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
    \]
    so we have well-defined-ness on the nose. 
    We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
    % \begin{itemize}
        % \item If there exists $z' = <_{c,Z}^{\mu}(z)$ then 
        % \[
        %     [<_{c,X+Y}^{\mu}(f_{E};i_{1,E}(z))] = [f_{E};i_{1,E}(<_{c,Z}^{\mu}(z))] = [f_{E};i_{1,E}(z')]
        % \]
        % \[
        %     [<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z))] = [g_{E};i_{2,E}(<_{c,Z}^{\mu}(z))] = [g_{E};i_{2,E}(z')]
        % \]
        % and
        % \[ 
        %     [f_{E};i_{1,E}(z')] = [g_{E};i_{2,E}(z')]
        % \]
        % by commutativity.
        % \item The other conditions satisfy functionality by definition.

@@

    and for verticies
    \begin{enumerate}

@@

        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of verticies and hyperedges.

@@

    $V$s and $E$s are sets of verticies and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.

@@

\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
\end{proposition}
\begin{proof}
    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textsf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
    $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \llbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
    We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
    We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.

    Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
    And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
    \[
        \scalebox{0.8}{
        \tikzfig{combinatorial_semantics/DPOI_square}
        }
    \]
    We then further note that if $l$ is a subterm of $g$ then the nodes in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent.
    This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
    The no-dangling condition is satisfied because we consider only mda-cospans.
    Therefore, there exists a pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
    Then necessarily verticies in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do in $\mathcal{G}$ so.
    The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
    Hence $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
    The rest follows by induction on the number of applied equations.
    Similar argument then applies to other rules in $\mathcal{S}$ as well.
\end{proof}

>>>>

\usepackage{subcaption}
\usepackage{caption}

@@

\begin{figure}
\[
	\scalebox{0.65}{
	\tikzfig{categorical-semantics/egraph-strings}
	}
\]
\captionsetup{belowskip=-5ex}
\caption{String diagrams for semilattice enriched symmetric monoidal categories.}
\label{fig:egraph-strings}
\end{figure}

@@

\begin{remark}
    Both $<$ and $\consistency$ can be considered as functions defined on $V_{\mathcal{F}} + E_{\mathcal{F}}$, \textit{i.e.}, on the coproduct of vertices and edges.
    To make things well-typed, we will use corresponding coproduct injections $\iota_{V} : {V_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ and $\iota_{E} : {E_{\mathcal{F}}} \to V_{\mathcal{F}} + E_{\mathcal{F}}$ when passing either a vertex or an edge into these functions.
    For example, an immediate successor of a vertex $x$ can be written functionally as $<_{\mathcal{F}}^{\mu}(\iota_{V_{\mathcal{F}}}(x))$.
\end{remark}

@@

We then quotient the set vertices and edges in $X + Y$ by $R_{V}$ and $R_{E}$ respectively.
    \begin{align*}
        \sfrac{X + Y}{\sim (R_{V},R_{E})} = \{&\\
            &\sfrac{V_{X} + V_{Y}}{\sim (R_{V},R_{E})}, \sfrac{E_{X} + E_{Y}}{\sim (R_{V},R_{E})}, \sfrac{s_{X+Y}}{\sim (R_{V},R_{E})},\\
            &\sfrac{t_{X,Y}}{\sim (R_{V},R_{E})}, \consistency_{\sfrac{X+Y}{\sim (R_{V},R_{E})}}, <_{\sfrac{X + Y}{\sim (R_{V},R_{E})}}\\
        & \}    
    \end{align*}
We will refer to $\sim (R_{V},R_{E})$ just as $\sim$, \textit{i.e.}, by writing $\sfrac{V_{X} + V_{Y}}{\sim}$ and concrete relation will be clear from the context.
We have 
\[
    \sfrac{s_{X+Y}}{\sim} : \sfrac{E_{X} + E_{Y}}{\sim} \to (\sfrac{V_{X} + V_{Y}}{\sim})^{*}
\]
and there is an obvious surjective function $[-]_{V} : (V_{X} + V_{Y}) \to (\sfrac{V_{X} + V_{Y}}{\sim})$ that maps elements to their equivalence classes and $[-]_{V}^{*}$ is its extension to sequences.
There is also $[-] : E_{X} + E_{Y} \to \sfrac{E_{X} + E_{Y}}{\sim}$ and we will omit subscripts as the correct type will be clear from the argument.
We then define 
\[
    \sfrac{s_{X+Y}}{\sim}([e]) = s_{X+Y};[-]^{*}(e) = [s_{X+Y}(e)]^{*}
\]
We will also use subscripts when it is important to tell if an element of $E_{X} + E_{Y}$ has a pre-image in either $E_{X}$ or $E_{Y}$ by writing $e_{x}$ or $e_{y}$.
Likewise, 
\[
    t_{\sfrac{X+Y}{\sim}}([e]) = [t_{X+Y}(e)]^{*}
\]

@@

Recall that $<$ is essentially a transitive clousure of $<^{\mu}$ and hence we can first define 
    \[<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]_{E})
    \] and 
    \[
        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]_{V})
    \] where $\iota_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
    We will consider several cases.
    First, assume that $u$ is an element of $V_{X+Y}$ and
    \begin{enumerate}
        % \item $u$ has a pre-image $z$ in $V_{Z}$, $[z) \not = \varnothing$.
        %       The last requirement means that there necessarily exist $x = <_{c,X}^{\mu}(f_{E}(z)) = f_{E}(<_{c,Z}^{\mu}(z))$ and $y = <_{c,Y}^{\mu}(g_{E}(z)) = g_{E}(<_{c,Z}^{\mu}(z))$.
        %       We then define $<_{c,\sfrac{X+Y}{\sim}}^{\mu}([u]) = [<_{c,X+Y}^{\mu}(u)]$
        \item $u$ has a pre-image $z$ in $V_{Z}$ (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) = \varnothing$, $[g_{V}(z)) \not = \varnothing$.
              The last condition means that there exists $e_y = <_{Y}^{\mu}(\iota_{V_{Y}}(g_{V}(z)))$.
              We then define 
              \[
                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [g_{V};\iota_{V_{Y}};<_{Y}^{\mu};\iota_{2,E}(z)] = [\iota_{2,E}(e_y)]
              \]
        \item $u$ has a pre-image $z$ in $V_{Z}$, (by discreteness, $[z) = \varnothing$) and $[f_{V}(z)) \not = \varnothing$ $[g_{V}(z)) = \varnothing$.
              The one before the last condition means that there exists $e_x = <_{X}^{\mu}(\iota_{V_{X}}(f_{V}(z)))$.
              We then define 
              \[
                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)] = [\iota_{1,E}(e_x)]
              \]
        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u)) = e'$
        \[
            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))] = [e']
        \]
        \item $u$ has no pre-image in $V_{Z}$ and $<_{X+Y}^{\mu}(\iota_{V_{X} + V_{Y}}(u))$ is undefined (\textit{i.e.}, $[u) = \varnothing$). 
              If there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}([v])) = [e']$ and such that there is an \textit{undirected} path from $[u]$ to $[v]$, then we define
              \[ 
                <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[v])
              \]
    \end{enumerate}
    % \question{In the case when images of $f$ and $g$ are top-level the function is undefined. Shall we introduce bottom?}
    Now, assume that $u$ is an element of $E_{X+Y}$.
    We have the only case of $u$ having no pre-image in $E_{Z}$ because it is empty.
    Consider two cases.
    \begin{enumerate}
    \item  $<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u)) = u_1$.
            Then we define
        \[
            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = [<_{X+Y}^{\mu}(\iota_{E_{X} + E_{Y}}(u))]
        \]
    \item $[u) = \varnothing$ and there exists $[v]$ such that $<_{\sfrac{X+Y}{\sim}}^{\mu}([v]) = [e']$ and such that there is an undirected path from $[u]$ to $[v]$, we define 
        \[
            <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{E_{X} + E_{Y}}{\sim}}[u]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[v])
        \] 
    \end{enumerate}
    % We will also build this relation in steps, first we do steps for vertices (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for vertices until fixed point.
    % Essentially, this interleaving means that vertices inherit the relation from incident edges and vice versa.
    Clearly, all the cases above are disjoint.

@@

    Let's check that this is a well-defined function.
    Suppose $t_1, t_2$ are vertices and $t_1 \sim t_2$, then $<_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X}+V_{Y}}{\sim}}[t_1]) = <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2])$.
    $t_1 \sim t_2$ implies there is $z$ such that $t_1 = f_{V};\iota_{1,V}(z)$ and $t_2 = g_{V};\iota_{2,V}(z)$ and two cases
    \begin{itemize}
        \item $[f_{V}(z)) \not = \varnothing$ and $[g_{V}(z)) = \varnothing$
        \item $[f_{V}(z)) = \varnothing$ and $[g_{V}(z)) \not = \varnothing$
    \end{itemize} 
    Consider the first case. Then, by definition.
    \[
        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_1]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
    \]
    and
    \[
        <_{\sfrac{X+Y}{\sim}}^{\mu}(\iota_{\sfrac{V_{X} + V_{Y}}{\sim}}[t_2]) = [f_{V};\iota_{V_{X}};<_{X}^{\mu};\iota_{1,E}(z)]
    \]
    so we have well-defined-ness on the nose. 
    We do not need to check well-defined-ness for edges as $e_1 \sim e_2$ just implies $e_1 = e_2$ for edges.
    % \begin{itemize}
        % \item If there exists $z' = <_{c,Z}^{\mu}(z)$ then 
        % \[
        %     [<_{c,X+Y}^{\mu}(f_{E};i_{1,E}(z))] = [f_{E};i_{1,E}(<_{c,Z}^{\mu}(z))] = [f_{E};i_{1,E}(z')]
        % \]
        % \[
        %     [<_{c,X+Y}^{\mu}(g_{E};i_{2,E}(z))] = [g_{E};i_{2,E}(<_{c,Z}^{\mu}(z))] = [g_{E};i_{2,E}(z')]
        % \]
        % and
        % \[ 
        %     [f_{E};i_{1,E}(z')] = [g_{E};i_{2,E}(z')]
        % \]
        % by commutativity.
        % \item The other conditions satisfy functionality by definition.

@@

    and for vertices
    \begin{enumerate}

@@

        The pushout in $EHyp_{\Sigma}$ should also be a pushout on the underlying sets of vertices and hyperedges.

@@

    $V$s and $E$s are sets of vertices and hyperedges respectively, marked squares are pushouts and $\mathcal{L}^{\bot}$s are pushout complements.

@@

\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
\end{proposition}
\begin{proof}
    To prove this statement we essentially need to show that the functor $\llbracket - \rrbracket$ is well-defined, \textit{i.e.}, if $g = h$ in $\textsf{PROP}^{+}(\Sigma)$ then $\llbracket g \rrbracket = \llbracket h \rrbracket$ in $\WellTypedMdaEcospans / \mathcal{S}$ or $\llbracket g \rrbracket \Rrightarrow_{\mathcal{S}}^{*} \llbracket h \rrbracket$ in $\WellTypedMdaEcospans$.
    $g = h$ implies that there is a sequence of SMC and enrichment equations such that when applied to $g$ it is turned into $h$.
    As SMC laws are subsumed by e-hypergraph representation we need to show that for each enrichment equation $l = r$ that turns $g$ into $h$ there is a pushout complement $\mathcal{C}$ such that $\llbracket g \rrbracket = \mathcal{C}[\llbracket l \rrbracket]$ and such that $\llbracket h \rrbracket = \mathcal{C}[\llbracket r \rrbracket]$, where $\mathcal{C}[-]$ is defined by computing the pushout along the common interfaces.
    We first note that any subterm of $g$ correspond to a convex down-closed subgraph of $\llbracket f \rrbracket$ and hence there is always a convex down-closed match of $\llbracket l \rrbracket$ within $\llbracket g \rrbracket$.
    We then need to show that for all $\llbracket l \rrbracket = \llbracket r \rrbracket$ in $\mathcal{S}$ the match satisfies the conditions in Proposition~\ref{prop:complement_existence} which will complete the first square of EDPOI and then check that the arrows in the second square satisfy the condition from Theorem~\ref{th:existence_of_pushouts}.

    Consider the rule $\llbracket l \rrbracket = \llbracket r \rrbracket$ in Figure~\ref{fig:structural_rule_pushout} which corresponds to the top-left equation in Figure~\ref{fig:string-equations}.
    And recall the diagram assuming $\mathcal{L} = \llbracket l \rrbracket$, $\mathcal{R} = \llbracket r \rrbracket$, $\mathcal{G} = \llbracket g \rrbracket$ and $\mathcal{H} = \llbracket h \rrbracket$.
    \[
        \scalebox{0.8}{
        \tikzfig{combinatorial_semantics/DPOI_square}
        }
    \]
    We then further note that if $l$ is a subterm of $g$ then the nodes in the image of input and output external interfaces of $\llbracket l \rrbracket$ in $\llbracket g \rrbracket$ all share the same set of parents and are pairwise consistent if the set of parents is non-empty.
    This satisfies conditions (1)-(2) in Proposition~\ref{prop:complement_existence}.
    The no-dangling condition is satisfied because we consider only mda-cospans.
    Therefore, there exists a pushout complement $\mathcal{C}$ which is $\llbracket g \rrbracket$ with $\llbracket l \rrbracket$ removed.
    Then necessarily vertices in $\mathcal{C}$ in the image of $i+j$ share the same set of parents and are pairwise consistent as they do so in $\mathcal{G}$.
    The same is true for the image of $i + j$ in $\mathcal{R}$ and therefore the necessary pushout exists which is $\mathcal{C}$ with $\mathcal{R} = \llbracket r \rrbracket$ glued in and it is equal to $\llbracket h \rrbracket$.
    Hence $\llbracket g \rrbracket \Rrightarrow{} \llbracket h \rrbracket$.
    The rest follows by induction on the number of applied equations.
    Similar argument then applies to other rules in $\mathcal{S}$ as well.
\end{proof}
